### 2020-01-30 ChapterRampInput/ChapterRampInput.tex
Preliminaries ready

\chapter{Ramp input estimation method}\label{chap:RampInput}

\begin{quote}
\emph{The results of this chapter ...}\vfill{}
\end{quote}

\section{Introduction}

@@

The affine input is modeled as a straight line $u = {a} t + {b}$ with parameters the slope $a$ and the interception $b$.
The affine input estimation problem is formulated as a signal processing problem as follows.

\paragraph{Problem} 
Given the sequence of measured output observations $\mathbf{y} = \left( y(1),\ldots,y(T)\right), y(t) \in {\rm I\!R}$, of a stable linear time-invariant system of order $n$, and static gain $\gamma$, generated by an affine input $u = \widebar{a} t + \widebar{b}$, estimate the parameters of the affine input, \textit{i.e.,} find the values of the parameters $\widehat{a}, \widehat{b} \in {\rm I\!R}$ such that $\widehat{u} = \widehat{a} t + \widehat{b}$ approximates $u$.
The measured observations $\mathbf{y} = \widebar{\mathbf{y}} + \bm{\epsilon}$ are exact sensor responses $\widebar{\mathbf{y}}$ perturbed by additive noise  $\bm{\epsilon}$ assumed to be independent and normally distributed of zero mean and given variance $\sigma_{\epsilon}^2$.

@@

RLS is a special case of the EWRLS that can solve the minimization problem (\ref{eqn:min_ls}).
The minimization problem is a structured EIV problem and its EWRLS solution is equivalent to 
\begin{equation} \widehat{\mathbf{x}} = \underset{\mathbf{x}}{\mathrm{argmin}} \ \left\Vert \mathbf{W}^{1/2} \left( \mathbf{y} - \mathbf{K} \mathbf{x} \right) \right\Vert^2_2 . \label{eqn:min_ewrls} \end{equation}
where $\mathbf{y}$, $\widehat{\mathbf{x}}$, and $\mathbf{K}$ are defined in (\ref{eqn:min_ls}) and (\ref{eqn:matrix}) and
$\mathbf{W} \in \mathbb{R}^{(T-n) \times (T-n)}$ is a diagonal matrix of descending powers of the weight $\lambda \in [0, 1)$, \textit{i.e.}, $W = \mathrm{diag}\left(\lambda^{T-n}, \ldots, \lambda^2, \lambda^1 \right)$.
The weight $\lambda$ is a data selection forgetting factor since it enables to apply different weights to the residuals $\mathbf{y} - \mathbf{K} \mathbf{x}$.
When $\lambda=1$, we have the same solution as RLS.
When $\lambda<1$, the older residuals are weighted with lower values than the residuals of recent observations.
In this way, the solution of the minimization problem depends more on new data.

The profile of an affine input excitation can be reconstructed by solving the structured errors-in-variables problem  (\ref{eqn:min_ewrls}) where $y$ is the corresponding response of a stable dynamic system, $\mathbf{K}$ is constructed from the output observations, and $\lambda<1$.

After the affine input $\widehat{u}$ has been estimated, the affine input parameters $a$ and $b$ are estimated by fitting $\widehat{u}$ to the straight line $\widehat{u} = a t + b$ using linear regression, as follows
\begin{equation} 
  \begin{bmatrix} 1 & 1  \\ \vdots & \vdots \\ T-\tau+1 & 1 \end{bmatrix} 
  \begin{bmatrix} \widehat{a} \\ \widehat{b}  \end{bmatrix} = 
  \begin{bmatrix} \widehat{u}(\tau) \\ \vdots \\ \widehat{u}(T) \end{bmatrix} \label{eqn:LS}
\end{equation}
where $\tau$ is a number of samples used as a tuning parameter that counteracts the time delay that exists when a LTI system is excited with an affine input $u$.
The subspace method can process online the sensor response to the affine excitation.
For each new observation $y(t)$, the estimation $\widehat{u}$ is updated followed by the update of the slope $\widehat{a}$ and the interception $\widehat{b}$.
The values of the tuning parameters $\lambda$ and $\tau$ can be obtained in the calibration of the method using the response of the sensor.

@@

\subsection{Maximum-likelihood method}
Using a model of the sensor, the ML method estimates the sensor initial conditions and the parameters of the applied affine input.
The affine input parameters $a$ and $b$ are the slope and the interception of the straight line model $u = at + b$.

Algorithm \ref{Alg1} lists the steps of the proposed ML method.
This is an iterative minimization method that uses the Jacobian of the residual error function $\mathbf{r} = \widehat{\mathbf{y}} - \mathbf{y}$ to search the direction that minimizes the difference between the measured sensor response $\mathbf{y}$ and the simulated response $\widehat{\mathbf{y}}$.

\begin{algorithm}
\caption{ML Affine input estimation.}\label{Alg1}
\begin{algorithmic}
  \REQUIRE{$\mathbf{y}$, and sensor model parameters}{}
  \STATE {Initialize $\mathbf{\theta} = (a, b, \mathbf{x}_{\text{ini}})$}{}
  \FOR{each $N$ observations of $\mathbf{y}$}{}
  \STATE {Simulate model response $\widehat{\mathbf{y}}(\mathbf{\theta})$}{}
  \STATE {Compute error $\mathbf{r}(\mathbf{\theta}) = \mathbf{y} - \widehat{\mathbf{y}}(\mathbf{\theta})$}{}
  \STATE {Minimize $\mathbf{r}^\top \mathbf{r}$ over $\mathbf{\theta}$}{}
  \STATE {\hspace{0.5cm}  using analytic Jacobian $\partial \mathbf{r} / \partial \mathbf{\theta}$}{}
  \STATE {Update $\mathbf{\theta}$}{}
\ENDFOR
\ENSURE {Optimized parameters $\widehat{a}, \widehat{b}$, and $\widehat{\mathbf{x}}_{\text{ini}}$}{}

@@

To initialize the optimization variables $a$, $b$, and $\mathbf{x}_{\text{ini}}$,
we use the subspace estimation method using at least the first $2n+2$ samples.
With the initial affine input parameters we simulate a sensor excitation and, since we are using few samples, we have an approximation of the initial conditions.
The optimization variables updates are computed every $N$ new observations.
The minimization can be done using the Levenberg-Marquardt algorithm \cite{Nocedal06}.

@@

The ML method simulates a dynamic system, and computes the Jacobian of the residual error in each iteration.
The analytic formulation of the Jacobian benefits the estimation method in two ways: it speeds up the minimization and gives direct access to the variance of the estimates.
The covariance matrix of the ML estimates can be expressed as \cite{Pintelon12Book}
\begin{equation} \mathbf{Cov} \left( \widehat{x} \right) = \left( \left( \dfrac{\partial \mathbf{r} }{ \partial \mathbf{\theta} } \right)^\top \left( \dfrac{\partial \mathbf{r} }{ \partial \mathbf{\theta} } \right) \right)^{-1}. \label{eqn:covOpt} \end{equation}

\subsubsection{ML affine input estimation example}
We use the previously described dynamic weighing system to illustrate the ML method.
In this case, the formulation of the problem is: 
 \begin{equation} \begin{aligned}
     & \text{Minimize} \quad \text{over} \ a, b, \mathbf{x}_{\text{ini}} \quad \mathbf{r}^T \mathbf{r} \text{, subject to:} \\ & \ \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 & \frac{x_{\text{ini,1}}}{a t + b + m} \\ g & \frac{x_{\text{ini,2}}}{a t + b + m}  \end{bmatrix} \begin{bmatrix} 1 \\ \delta(t)  \end{bmatrix}, \\ & \ \widehat{y} = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} .  
 \end{aligned} \end{equation}
where $a$, $b$ are the affine input parameters, $m$, $k_{\mathrm{d}}$, and $k_{\mathrm{s}}$ are the model parameters, the model states $\mathbf{x}$ are the position and the speed of the weighing scale, the difference between the sensor response $\mathbf{y}$ and the simulated response $\widehat{\mathbf{y}}$ is the residual $\mathbf{r} = \widehat{\mathbf{y}} - \mathbf{y}$.  
The model initial conditions $\mathbf{x}_{\text{ini}}$ are considered optimization variables
and appear in the augmented column of the input matrix.

@@

\subsection{Time-varying compensation filter}
The time-varying (TV) filter described in \cite{Pietrzak14} was designed to compensate the measured responses of a conveyor weighing system, considering they are modeled as a saturated ramp.
The TV filter consists of three low-pass infinite impulse response (IIR) filters in cascade, where the $i-\mathrm{th}$ IIR filter is given by
\begin{equation} \widehat{y}_i(t) + k_1(t) \widehat{y}_i(t-1) = k_2(t) \left( \widehat{y}_{i-1}(t) + \widehat{y}_{i-1}(t-1) \right) \end{equation}
for $i = 1,\ldots,3$ and $t=0,\ldots,T$.
The sensor response is fed to the filter, then $\widehat{y}_0(t) = y(t)$, and the output of the TV filter $\widehat{u}_\mathrm{ltv}(t) = \widehat{y}_3(t)$ is an estimation of the affine input.
Since in our case we are processing only the ramp without the saturation, the estimates $\widehat{a}_\mathrm{ltv}$ and $\widehat{b}_\mathrm{ltv}$ of the input parameters are obtained by fitting a straight line to the estimated input $\widehat{u}_\mathrm{ltv}$ using linear regression.

The time-varying coefficients $k_1(t)$ and $k_2(t)$ are computed from
\begin{equation} k_1(t) = \frac{f_c(t) - \frac{k_3}{\pi T_s}}{f_c(t) + \frac{k_3}{\pi T_s}}, \quad k_2(t) = \frac{1 + k_1(t)}{2} , \quad k_3 = \sqrt{ \sqrt[3]{2}  - 1} \end{equation}
where $T_s$ is the sampling time and $f_c(t)$ is a heuristic "cutoff" frequency
\begin{equation} f_c(t) = f_u + \left( f_l - f_u \right) \beta^{\left( t-1 \right) /  \alpha \left( T-1 \right) } \end{equation}
that changes between the lower $f_l$ and upper $f_u$ limits, where the coefficient $\beta$ is lower than one, and $\alpha$ is the decay rate. 
The lower frequency value $f_l$ and the coefficient $\beta$ are fixed and the variables $f_u$ and $\alpha$ are optimized off-line by solving the minimization problem 
\begin{equation} \mathrm{minimize} \ \quad \ \mathrm{over} \ f_u, \alpha \ \quad \ \mathrm{max} \left( \dfrac{\mu_{\widetilde{a}_{\mathrm{ltv}}}}{\mu_{\mathrm{spec}}}, \dfrac{\mu_{\widetilde{b}_{\mathrm{ltv}}}}{\mu_{\mathrm{spec}}}, \dfrac{\sigma_{\widetilde{a}_{\mathrm{ltv}}}}{\sigma_{\mathrm{spec}}}, \dfrac{\sigma_{\widetilde{b}_{\mathrm{ltv}}}}{\sigma_{\mathrm{spec}}} \right) + \mathrm{max} \left( \dfrac{\eta_{i}}{T} \right) \label{eqn:tv_optim} \end{equation}
where $\mu_{\widetilde{a}_\mathrm{ltv}}$, $\mu_{\widetilde{b}_\mathrm{ltv}}$, $\sigma_{\widetilde{a}_\mathrm{ltv}}$ and $\sigma_{\widetilde{b}_\mathrm{ltv}}$ are the mean values and the standard deviations of the estimation errors $\widetilde{a}_\mathrm{ltv} = \widehat{a}_\mathrm{ltv} - \widebar{a}$, and $\widetilde{a}_\mathrm{ltv} = \widehat{b}_\mathrm{ltv} - \widebar{b}$, and where $\widebar{a}$ and $\widebar{b}$ are the true values of the input parameters.
The values $\mu_{\mathrm{spec}}$ and $\sigma_{\mathrm{spec}}$ are specified in the OIML recommendation R51 \cite{OIML_R51_1} for mass measurements that use a conveyor belt.

@@

A second order weighing system was excited with an affine input to get the transient response.
The parameters of the weighing system are $m = 15$ g, $d = 5.5$ Ns/m and $k = 10250$ N/m.
The applied affine input $u = 100 t + 10$ represents a mass that changes from $10$ g to $110$ g, at a constant rate, in a time interval of $0.1$ s.
This change of mass represents one example of the weighing input in a conveyor weighing system when an object of 100 g is measured while it is moving at constant speed.
The sensor response is acquired using sampling time $T_s = 0.1$ ms.
In Figure \ref{fig:sensor_weight} the input $u$ is represented with the dotted line, the oscillatory curve is the corresponding sensor ramp response $y$, and $\widehat{u}_{\mathrm{}}$ is a typical input estimate obtained with the subspace method.

>>>>

\chapter{Affine input estimation method}\label{chap:AffineInput}

\begin{quote}
\emph{The results of this chapter were published in Quintana Carapia, G., and Markovsky, I.,
"Input parameters estimation from time-varying measurements", Measurement Journal, Vol. 153, 2020,
ISSN 0263-2241, doi:10.1016/j.measurement.2019.107418. \nocite{QuintanaMEAS} }\vfill{}
\end{quote}

\vfill{}

%\section{Introduction}

@@

The affine input is modeled as a straight line $u(t) = {a} t + {b}$ with parameters the slope $a$ and the interception $b$.
The affine input estimation problem is formulated as a signal processing problem as follows.

\paragraph{Problem} 
Given the sequence of measured output observations $\widetilde{\mathbf{y}} = \left( \widetilde{y}(1),\ldots, \widetilde{y}(T)\right)$, with $\widetilde{y}(t) \in {\rm I\!R}$, of a stable linear time-invariant system of order $n$, and static gain $\gamma$, generated by an affine input $u(t) = \widebar{a} t + \widebar{b}$, estimate the parameters of the affine input, \textit{i.e.,} find the values of the parameters $\widehat{a}, \widehat{b} \in {\rm I\!R}$ such that $\widehat{u}(t) = \widehat{a} t + \widehat{b}$ approximates $u$(t).
The measured observations $\widetilde{\mathbf{y}} = \widebar{\mathbf{y}} + \bm{\epsilon}$ are exact sensor responses $\widebar{\mathbf{y}}$ perturbed by additive noise  $\bm{\epsilon}$ assumed to be independent and normally distributed of zero mean and given variance $\sigma_{\epsilon}^2$.

@@

RLS is a special case of the EWRLS that can solve the errors-in-variables (EIV) minimization problem (\ref{eqn:min_seiv}).
The minimization problem is a structured EIV problem and its EWRLS solution is equivalent to 
\begin{equation} \widehat{\mathbf{x}} = \underset{\mathbf{x}}{\mathrm{argmin}} \ \left\Vert \mathbf{W}^{1/2} \left( \widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \mathbf{x} \right) \right\Vert^2_2 . \label{eqn:min_ewrls} \end{equation}
where $\widetilde{\mathbf{y}}$, $\widehat{\mathbf{x}}$, and $\widetilde{\mathbf{K}}$ are defined in (\ref{eqn:min_seiv}) and (\ref{eqn:K0plusnoise}) and
$\mathbf{W} \in \mathbb{R}^{(T-n) \times (T-n)}$ is a diagonal matrix of descending powers of the weight $\lambda \in [0, 1)$, \textit{i.e.}, $\mathbf{W} = \mathrm{diag}\left(\lambda^{T-n}, \ldots, \lambda^2, \lambda^1 \right)$.
The weight $\lambda$ is a data selection forgetting factor since it enables to apply different weights to the residuals $\widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \mathbf{x}$.
When $\lambda=1$, we have the same solution as RLS.
When $\lambda<1$, the older residuals are weighted with lower values than the residuals of recent observations.
In this way, the solution of the minimization problem depends more on new data.

The profile of an affine input excitation can be reconstructed by solving the structured errors-in-variables problem  (\ref{eqn:min_ewrls}) where $y$ is the corresponding response of a stable dynamic system, $\widetilde{\mathbf{K}}$ is constructed from the output observations, and $\lambda<1$.

After the affine input $\widehat{\mathbf{u}}$ has been estimated, the affine input parameters $a$ and $b$ are estimated by fitting $\widehat{\mathbf{u}}$ to a straight line using linear regression, as follows
\begin{equation} 
  \begin{bmatrix} 1 & 1  \\ \vdots & \vdots \\ T-\tau+1 & 1 \end{bmatrix} 
  \begin{bmatrix} \widehat{a} \\ \widehat{b}  \end{bmatrix} = 
  \begin{bmatrix} \widehat{u}(\tau) \\ \vdots \\ \widehat{u}(T) \end{bmatrix} \label{eqn:LS}
\end{equation}
where $\tau$ is a number of samples used as a tuning parameter that counteracts the time delay that exists when a LTI system is excited with an affine input $\mathbf{u}$.
The subspace method can process online the sensor response to the affine excitation.
For each new observation $y(t)$, the estimation $\widehat{\mathbf{u}}$ is updated followed by the update of the slope $\widehat{a}$ and the interception $\widehat{b}$.
The values of the tuning parameters $\lambda$ and $\tau$ can be obtained in the calibration of the method using the response of the sensor.

@@

\subsubsection{Statistical analysis of the subspace method}

To obtain the first and second moments of the step input estimate $\widehat{\mathbf{u}}$, we need to study the solution
\begin{equation} \widehat{\mathbf{x}} = ( \widetilde{\mathbf{K}}^\top \mathbf{W} \widetilde{\mathbf{K}}  )^{-1} \widetilde{\mathbf{K}}^\top \mathbf{W} \widetilde{\mathbf{y}} , \label{eqn:xhatW} \end{equation}
of the overdetermined structured errors-in-variables EIV problem (\ref{eqn:min_ewrls}).
Using a second order Taylor series expansion of the inverse matrix we can approximate the LS solution as
\begin{equation} \widehat{\mathbf{x}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{C}^{-1} (\mathbf{K} + \mathbf{E})^\top \mathbf{W} (\mathbf{y} + \bm{\epsilon}). \label{eqn:xhatexpW} \end{equation} 
where 
\begin{equation} \mathbf{C} = \mathbf{K}^\top \mathbf{W} \mathbf{K}, \quad \text{and} \quad \mathbf{M} = \mathbf{C}^{-1} ( \mathbf{K}^\top \mathbf{W} \mathbf{E} + \mathbf{E}^\top \mathbf{W} \mathbf{K} + \mathbf{E}^\top \mathbf{W} \mathbf{E} ). \end{equation}

The Taylor series approximation of $\widehat{\mathbf{x}}$ enables the calculation of the estimate bias and covariance since the measurement noise $\bm{\epsilon}$ and $\mathbf{E}$ are no more subject to matrix inversion. 
The bias and the covariance of the estimate $\widehat{\mathbf{x}}$ are obtained from the definitions (\ref{eqn:biasdef}), and (\ref{eqn:covdef}). %The bias and the covariance of the estimate $\widehat{\mathbf{x}}$ are obtained from \begin{equation}  \mathbf{b} \left(\widehat{\mathbf{x}} \right) = \mathbf{\mu} - \widebar{\mathbf{x}}, \label{eqn:biasdef} \end{equation}  \begin{equation} \begin{aligned} \mathrm{\mathbf{Cov}} \left( \widehat{\mathbf{x}} \right) & = \mathbb{E} \left\{ \left( \widehat{\mathbf{x}} - \mathbf{\mu} \right)  \left( \widehat{\mathbf{x}} - \mathbf{\mu} \right)^\top \right\} . \end{aligned} \label{eqn:covdef} \end{equation} where $\mathbf{\mu} = \mathbb{E} \left\{ \widehat{\mathbf{x}} \right\}$ is the expected value of the estimate and $\widebar{\mathbf{x}}$ is the true value. 
Considering the structure of the EIV problem, the bias and the covariance of the estimate approximation (\ref{eqn:xhatexpW}) can be expressed as
\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{C}^{-1} \left(  \left( \mathbf{K}^\top \mathbf{W} \mathbf{B}_1 - \mathbf{B}_2 \right) \mathbf{x} - \left( \mathbf{K}^\top \mathbf{W} \mathbf{B}_3 - \mathbf{B}_4 \right) \right), \end{aligned} \label{eqn:biasEW} \end{equation}
\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \mathbf{K}^\dagger \mathbf{W} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \mathbf{W} \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{x} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{x} \right), \end{aligned} \label{eqn:varEW} \end{equation}
where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{W} \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \mathbf{W} \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \mathbf{x}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{x} \bm{\epsilon}^\top \Big\}$, $\mathbf{P}_\perp = \mathbf{I} - \mathbf{K} \mathbf{K}^\dagger \mathbf{W}$, and $\mathbf{K}^\dagger$ is the pseudo-inverse matrix of $\mathbf{K}$.

The bias and covariance given by expressions (\ref{eqn:biasEW}) and (\ref{eqn:varEW}) depend on the unobservable true values $\mathbf{x}$ and $\mathbf{K}$.
The measured observations are in the sensor step response $\widetilde{\mathbf{y}}$, and from its observations we construct $\widetilde{\mathbf{K}}$ and compute $\widehat{\mathbf{x}}$.
The substitution of the measured data in the expressions gives an approximation of the estimate bias and covariance. 
We have then
\begin{equation} \begin{aligned} \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \widetilde{\mathbf{C}}^{-1} \left(  \left( \widetilde{\mathbf{K}}^\top \mathbf{W} \widetilde{\mathbf{B}}_1 - \widetilde{\mathbf{B}}_2 \right) \widehat{\mathbf{x}} - \left( \widetilde{\mathbf{K}}^\top \mathbf{W} \widetilde{\mathbf{B}}_3 - \widetilde{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasSTW} \end{equation}
  \begin{equation} \begin{aligned} \widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \mathbf{W} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \mathbf{W} \widetilde{\mathbf{K}}^{\dagger \top} - \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{x} \right) \widetilde{\mathbf{b}}_{\mathrm{p}}^\top \left( \widehat{x} \right), \end{aligned} \label{eqn:varSTW} \end{equation}
  where $\widetilde{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \mathbf{W} \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \widetilde{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \mathbf{W} \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \mathbf{W} \widetilde{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{C}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \widehat{\mathbf{x}}^\top \mathbf{E}^\top \Big\}$, $\widetilde{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\mathbf{x}} \bm{\epsilon}^\top \Big\}$, and $\widetilde{\mathbf{P}}_\perp = \mathbf{I} - \widetilde{\mathbf{K}} \widetilde{\mathbf{K}}^\dagger \mathbf{W}$.

The results of the expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$ can be found using Lemma \ref{lem:lemma1}.
The bias and covariance were obtained to extend previous analysis conducted on EIV estimation problems without an imposed structure \citep{Vaccaro94, Stewart90SPT}.
It was shown that the bias and variance expressions (\ref{eqn:biasSTW}) and (\ref{eqn:varSTW}) are valid predictions of the first and second moments of the LS estimate of a Hankel structured EIV problem.
The problem formulated by the step input estimation method belongs to this type of structured EIV problems and can use the derived expressions to find the bias and variance of the input estimate $\widehat{u}$.
The bias of the estimate $\widehat{\mathbf{u}}$ is the fist element of $\widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right)$ and the variance of $\widehat{\mathbf{u}}$ is the first element in the main diagonal of $\widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\mathbf{x}} \right)$.

\subsection{Maximum-likelihood method}
Using a model of the sensor, the ML method estimates the sensor initial conditions and the parameters of the applied affine input.
The affine input parameters $a$ and $b$ are the slope and the interception of the straight line model $u(t) = at + b$.

Algorithm \ref{Alg1} lists the steps of the proposed ML method.
This is an iterative minimization method that uses the Jacobian of the residual error function $\mathbf{r} = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}$ to search the direction that minimizes the difference between the measured sensor response $\widetilde{\mathbf{y}}$ and the simulated response $\widehat{\mathbf{y}}$.

\begin{algorithm}
\caption{ML Affine input estimation.}\label{Alg1}
\begin{algorithmic}
  \REQUIRE{$\widetilde{\mathbf{y}}$, and sensor model parameters}{}
  \STATE {Initialize $\mathbf{\theta} = (a, b, \mathbf{x}_{\text{ini}})$}{}
  \FOR{each $N$ observations of $\widetilde{\mathbf{y}}$}{}
  \STATE {Simulate model response $\widehat{\mathbf{y}}(\mathbf{\theta})$}{}
  \STATE {Compute error $\mathbf{r}(\mathbf{\theta}) = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}(\mathbf{\theta})$}{}
  \STATE {Minimize $\mathbf{r}^\top \mathbf{r}$ over $\mathbf{\theta}$}{}
  \STATE {\hspace{0.5cm}  using analytic Jacobian $\partial \mathbf{r} / \partial \mathbf{\theta}$}{}
  \STATE {Update $\mathbf{\theta}$}{}
\ENDFOR
\ENSURE {Optimized parameters $\widehat{a}, \widehat{b}$, and $\widehat{\mathbf{x}}_{\text{ini}}$}{}

@@

To initialize the optimization variables $a$, $b$, and $\mathbf{x}_{\text{ini}}$,
we use the subspace estimation method using at least the first $2n+2$ samples.
With the initial affine input parameters we simulate a sensor excitation and, since we are using few samples, we have an approximation of the initial conditions.
The optimization variables updates are computed every $N$ new observations.
The minimization can be done using the Levenberg-Marquardt algorithm \citep{Nocedal06}.

@@

The ML method simulates a dynamic system, and computes the Jacobian of the residual error in each iteration.
The analytic formulation of the Jacobian benefits the estimation method in two ways: it speeds up the minimization and gives direct access to the variance of the estimates.
The covariance matrix of the ML estimates can be expressed as \citep{Pintelon12Book}
\begin{equation} \mathbf{Cov} \left( \widehat{x} \right) = \left( \left( \dfrac{\partial \mathbf{r} }{ \partial \mathbf{\theta} } \right)^\top \left( \dfrac{\partial \mathbf{r} }{ \partial \mathbf{\theta} } \right) \right)^{-1}. \label{eqn:covOpt} \end{equation}

\subsubsection{ML affine input estimation example}
We use the previously described dynamic weighing system to illustrate the ML method.
In this case, the formulation of the problem is: 
 \begin{equation} \begin{aligned}
     & \text{Minimize} \quad \text{over} \ a, b, \mathbf{x}_{\text{ini}} \quad \mathbf{r}^T \mathbf{r} \text{, subject to:} \\ & \ \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \frac{-k_{\mathrm{s}}}{a t + b + m} & \frac{-(a + k_{\mathrm{d}})}{a t + b + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 & \frac{x_{\text{ini,1}}}{a t + b + m} \\ g & \frac{x_{\text{ini,2}}}{a t + b + m}  \end{bmatrix} \begin{bmatrix} 1 \\ \delta(t)  \end{bmatrix}, \\ & \ \widehat{\mathbf{y}} = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} .  
 \end{aligned} \end{equation}
where $a$, $b$ are the affine input parameters, $m$, $k_{\mathrm{d}}$, and $k_{\mathrm{s}}$ are the model parameters, the model states $\mathbf{x}$ are the position and the speed of the weighing scale, the difference between the sensor response $\widetilde{\mathbf{y}}$ and the simulated response $\widehat{\mathbf{y}}$ is the residual $\mathbf{r} = \widetilde{\mathbf{y}} - \widehat{\mathbf{y}}$.  
The model initial conditions $\mathbf{x}_{\text{ini}}$ are considered optimization variables
and appear in the augmented column of the input matrix.

@@

\subsection{Time-varying compensation filter}
The time-varying (TV) filter described in \citep{Pietrzak14} was designed to compensate the measured responses of a conveyor weighing system, considering they are modeled as a saturated ramp.
The TV filter consists of three low-pass infinite impulse response (IIR) filters in cascade, where the $i-\mathrm{th}$ IIR filter is given by
\begin{equation} \widehat{y}_i(t) + k_1(t) \widehat{y}_i(t-1) = k_2(t) \left( \widehat{y}_{i-1}(t) + \widehat{y}_{i-1}(t-1) \right) \end{equation}
for $i = 1,\ldots,3$ and $t=0,\ldots,T$.
The sensor response is fed to the filter, then $\widehat{y}_0(t) = \widetilde{y}(t)$, and the output of the TV filter $\widehat{u}_\mathrm{ltv}(t) = \widehat{y}_3(t)$ is an estimation of the affine input.
Since in our case we are processing only the ramp without the saturation, the estimates $\widehat{a}_\mathrm{ltv}$ and $\widehat{b}_\mathrm{ltv}$ of the input parameters are obtained by fitting a straight line to the estimated input $\widehat{u}_\mathrm{ltv}$ using linear regression.

The time-varying coefficients $k_1(t)$ and $k_2(t)$ are computed from
\begin{equation} k_1(t) = \frac{f_c(t) - \frac{k_3}{\pi T_s}}{f_c(t) + \frac{k_3}{\pi T_s}}, \quad k_2(t) = \frac{1 + k_1(t)}{2} , \quad k_3 = \sqrt{ \sqrt[3]{2}  - 1} \end{equation}
where $T_s$ is the sampling time and $f_c(t)$ is a heuristic "cutoff" frequency
\begin{equation} f_c(t) = f_u + \left( f_l - f_u \right) \beta^{\left( t-1 \right) /  \alpha \left( T-1 \right) } \end{equation}
that changes between the lower $f_l$ and upper $f_u$ limits, where the coefficient $\beta$ is lower than one, and $\alpha$ is the decay rate. 
The lower frequency value $f_l$ and the coefficient $\beta$ are fixed and the variables $f_u$ and $\alpha$ are optimized off-line by solving the minimization problem 
\begin{equation} \mathrm{minimize} \ \quad \ \mathrm{over} \ f_u, \alpha \ \quad \ \mathrm{max} \left( \dfrac{\mu_{\widetilde{a}_{\mathrm{ltv}}}}{\mu_{\mathrm{spec}}}, \dfrac{\mu_{\widetilde{b}_{\mathrm{ltv}}}}{\mu_{\mathrm{spec}}}, \dfrac{\sigma_{\widetilde{a}_{\mathrm{ltv}}}}{\sigma_{\mathrm{spec}}}, \dfrac{\sigma_{\widetilde{b}_{\mathrm{ltv}}}}{\sigma_{\mathrm{spec}}} \right) + \mathrm{max} \left( \dfrac{\eta_{i}}{T} \right) \label{eqn:tv_optim} \end{equation}
where $\mu_{\widetilde{a}_\mathrm{ltv}}$, $\mu_{\widetilde{b}_\mathrm{ltv}}$, $\sigma_{\widetilde{a}_\mathrm{ltv}}$ and $\sigma_{\widetilde{b}_\mathrm{ltv}}$ are the mean values and the standard deviations of the estimation errors $\widetilde{a}_\mathrm{ltv} = \widehat{a}_\mathrm{ltv} - \widebar{a}$, and $\widetilde{a}_\mathrm{ltv} = \widehat{b}_\mathrm{ltv} - \widebar{b}$, and where $\widebar{a}$ and $\widebar{b}$ are the true values of the input parameters.
The values $\mu_{\mathrm{spec}}$ and $\sigma_{\mathrm{spec}}$ are specified in the OIML recommendation R51 \citep{OIML_R51_1} for mass measurements that use a conveyor belt.

@@

A second order weighing system was excited with an affine input to get the transient response.
The parameters of the weighing system are $m = 15$ g, $d = 5.5$ Ns/m and $k = 10250$ N/m.
The applied affine input $u(t) = 100 t + 10$ represents a mass that changes from $10$ g to $110$ g, at a constant rate, in a time interval of $0.1$ s.
This change of mass represents one example of the weighing input in a conveyor weighing system when an object of 100 g is measured while it is moving at constant speed.
The sensor response is acquired using sampling time $T_s = 0.1$ ms.
In Figure \ref{fig:sensor_weight} the input $u$ is represented with the dotted line, the oscillatory curve is the corresponding sensor ramp response $\widetilde{\mathbf{y}}$, and $\widehat{\mathbf{u}}_{\mathrm{}}$ is a typical input estimate obtained with the subspace method.
